---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bespatial

  <!-- badges: start -->
  [![R-CMD-check](https://github.com/Nowosad/bespatial/workflows/R-CMD-check/badge.svg)](https://github.com/Nowosad/bespatial/actions)
  [![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
  [![Codecov test coverage](https://codecov.io/gh/Nowosad/bespatial/branch/master/graph/badge.svg)](https://app.codecov.io/gh/Nowosad/bespatial?branch=master)
  <!-- badges: end -->

The goal of **bespatial** is to calculate several entropy metrics for spatial data inspired by Boltzmann's entropy formula.
It includes metrics introduced by Cushman for landscape mosaics (Cushman (2015)), landscape gradients and point patterns (Cushman (2021)); by Gao et al. for landscape gradients (2017, 2019).

## Installation

<!-- You can install the released version of bespatial from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("bespatial") -->
<!-- ``` -->

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("Nowosad/bespatial")
```

## Example

```{r}
library(terra)
library(bespatial)
mosaic = rast(system.file("raster/mosaic.tif", package = "bespatial"))
ce1 = bes_m_cushman(mosaic, 1000)
plot(mosaic, main = round(ce1$value, 2))
```

<!-- ## Documentation -->

## References

- Cushman, S. A. (2015). Calculating the configurational entropy of a landscape mosaic. In Landscape Ecology (Vol. 31, Issue 3, pp. 481–489). Springer Science and Business Media LLC. https://doi.org/10.1007/s10980-015-0305-2 
- Cushman, S. (2018). Calculation of Configurational Entropy in Complex Landscapes. In Entropy (Vol. 20, Issue 4, p. 298). MDPI AG. https://doi.org/10.3390/e20040298 
- Cushman, S. A. (2021). Generalizing Boltzmann Configurational Entropy to Surfaces, Point Patterns and Landscape Mosaics. In Entropy (Vol. 23, Issue 12, p. 1616). MDPI AG. https://doi.org/10.3390/e23121616 
- Gao, Peichao, Hong Zhang, and Zhilin Li. "A hierarchy-based solution to calculate the configurational entropy of landscape gradients." Landscape Ecology 32(6) (2017): 1133-1146.
- Gao, Peichao and Zhilin Li. "Aggregation-based method for computing absolute Boltzmann entropy of landscape gradient with full thermodynamic consistency." Landscape Ecology (2019).

## Contribution

Contributions to this package are welcome - let us know if you have any suggestions or spotted a bug. 
The preferred method of contribution is through a GitHub pull request. 
Feel also free to contact us by creating [an issue](https://github.com/nowosad/bespatial/issues).
